systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2006-02-19 paper/trunk/systemfit_estimation.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimation}\label{sec:Estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider a system of $G$ equations, where the $i$th equation is of
the form 
\begin{equation}
   y_{i} = X_i \beta_i + u_i, \quad i = 1, 2, \ldots, G
\end{equation}
where $y_i$ is a vector of the dependent variable,
$X_i$ is a matrix of the exogenous variables,
$\beta_i$ is the coefficient vector and
$u_i$ is a vector of the disturbance terms of the $i$th equation.

We can write the °stacked° system as
\begin{equation}
   \left[ \begin{array}{c}
      y_1 \\ y_2\\ \vdots\\ y_G
   \end{array} \right] = 
   \left[ \begin{array}{cccc}
      X_1 & 0 & \cdots & 0\\
      0 & X_2 & \cdots & 0\\
      \vdots & \vdots & \ddots & \vdots\\
      0 & 0 & \cdots & X_G
   \end{array}\right] \cdot 
   \left[ \begin{array}{c}
      \beta_1 \\ \beta_2 \\ \vdots\\ \beta_G
   \end{array} \right] +
   \left[ \begin{array}{c}
      u_1 \\ u_2 \\ \vdots\\ u_G
   \end{array} \right]
\end{equation}

or more simply as
\begin{equation}
   y = X \beta + u 
\end{equation}

We assume that there is no correlation of the disturbance terms 
across observations:
\begin{equation}
   E \left( u_{it} \, u_{jt^*} \right) = 0 
   \; \forall \; t \neq t^*
\end{equation}
where $i$ and $j$ indicate the equation number 
and $t$ and $t^*$ denote the observation number.

However, we explicitly allow for contemporaneous correlation:
\begin{equation}
   E \left( u_{it} \, u_{jt} \right) = \sigma_{ij}
\end{equation}

Thus, the covariance matrix of the total system is
\begin{equation}
   E \left( u \, u' \right) = \Omega = \Sigma \otimes I
\end{equation}
where $\Sigma = \left[ \sigma_{ij} \right]$ is the residual covariance matrix
and $I$ is an identity matrix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ordinary least squares (OLS)}
\label{sec:ols}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted least squares (WLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Seemingly unrelated regression (SUR)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Two-Stage least squares (2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weighted two-stage least squares (W2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Three-Stage least squares (3SLS)}
\label{sec:3sls}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation}\label{sec:Estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Ordinary least squares (OLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted least squares (WLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Seemingly unrelated regression (SUR)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Two-Stage least squares (2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Weighted two-stage least squares (W2SLS)}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Three-Stage least squares (3SLS)}

### 2006-02-19 paper/trunk/systemfit_intro.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

This paper is organized as follows:
In the following Section~\ref{sec:Estimation} we
introduce the mathematics of estimating equation systems.
Section~\ref{sec:Restrictions} shows how linear restrictions can be
imposed.
Some other relevant issues about estimation of equation systems
are discussed in Section~\ref{sec:Other}.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit}, especially how the features presented in the previous
sections can be used.
Finally, a summary and outlook are presented in
Section~\ref{sec:Summmary}.

>>>>

This paper is organized as follows:
In the following Section~\ref{sec:statistics} we
introduce the statistical background of estimating equation systems.
Section~\ref{sec:Usage} demonstrates how to run
\pkg{systemfit} and especially how the features presented in the previous
section can be used.
Finally, a summary and outlook are presented in
Section~\ref{sec:Summmary}.

### 2006-02-19 paper/trunk/systemfit_summary.tex
rorganized the sections and subsections: there now a section called "Statistical Background" that includes the subsections (former sections) "Estimation", "Restrictions" and "Other issues" (a section of the "statistical background" is suggested by the JSS
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@614 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\paragraph{Unbalanced datasets}
Currently, the \pkg{systemfit} package requires
that all equations have the same number of observations.
However, many data sets have unbalanced observations.%
\footnote{
For instance,
forestry datasets typically contain observations of many inexpensive
observations (stem diameter) and few expensive observations such
as stem height.
}
Simply dropping data points
that do not contain observations for all equations
may considerably reduce the number of observations,
and thus, the information utilized in the estimation.
Hence, it is our intention to include the capability for estimations
with unbalanced data sets as described in \citet{schmidt77}
in future releases of \pkg{systemfit}.

\paragraph{Serial correlation and heteroscedasticity}
For all of the methods developed in the package, the disturbances of
the individual equations are assumed to be independent and identically
distributed (iid).
The package could be enhanced by the inclusion of methods to fit
equations with serially correlated and heteroscedastic disturbances.

\paragraph{Estimation methods}
We wish to include more sophisticated estimation methods such
as limited-information maximum likelihood (LIML),
full-information maximum likelihood (FIML),
and generalized methods of moments (GMM), in future releases.

\paragraph{Non-linear estimation}
The \pkg{systemfit} package provides a function to estimate
systems of non-linear estimations.
However, the function \code{nlsystemfit} is currently under development
and the results are not yet always reliable
due to convergence problems.
We hope to improve this behavior soon.

>>>>

\subsubsection*{Unbalanced datasets}
Currently, the \pkg{systemfit} package requires
that all equations have the same number of observations.
However, many data sets have unbalanced observations.%
\footnote{
For instance,
forestry datasets typically contain observations of many inexpensive
observations (stem diameter) and few expensive observations such
as stem height.
}
Simply dropping data points
that do not contain observations for all equations
may considerably reduce the number of observations,
and thus, the information utilized in the estimation.
Hence, it is our intention to include the capability for estimations
with unbalanced data sets as described in \citet{schmidt77}
in future releases of \pkg{systemfit}.

\subsubsection*{Serial correlation and heteroscedasticity}
For all of the methods developed in the package, the disturbances of
the individual equations are assumed to be independent and identically
distributed (iid).
The package could be enhanced by the inclusion of methods to fit
equations with serially correlated and heteroscedastic disturbances.

\subsubsection*{Estimation methods}
We wish to include more sophisticated estimation methods such
as limited-information maximum likelihood (LIML),
full-information maximum likelihood (FIML),
and generalized methods of moments (GMM), in future releases.

\subsubsection*{Non-linear estimation}
The \pkg{systemfit} package provides a function to estimate
systems of non-linear estimations.
However, the function \code{nlsystemfit} is currently under development
and the results are not yet always reliable
due to convergence problems.
We hope to improve this behavior soon.
